### 2022-02-09 index.Rmd
fix lix to doc from index

---
title: "*Please Participate in Review Important <br/> Amberson Towers Condominium Documents*"
author: "A Request from the Governance Committee <br/> of the Amberson Towers Condominium Council"
date: "2/8/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

***

### Please read these documents and comment

This document available in two formats. Choose the format you prefer:

* [*As an Adobe PDF document*](DraftRules.html)

* *As a web page*