### 2022-02-10 tuftetest.Rmd

---
title: "A Tufte Handout Example"
runningheader: "Tufte Handout with R Markdown" # only for pdf output
subtitle: "An implementation in R Markdown" # only for html output
author: "JJ Allaire and Yihui Xie"
date: "`r Sys.Date()`"
output:
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_html: default
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
bibliography: skeleton.bib
link-citations: yes
---

```{r setup, include=FALSE}
library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
```

## ARTICLE I {-}

This Declaration is prepared in accordance with the provisions of the Unit Property Act of the Commonwealth of Pennsylvania (Act of July 3, 1963, P.L. No. 196) for the purpose of submitting to the provisions of said Act the property described in Article II below.

## ARTICLE II {-}

ALL THAT CERTAIN LOT OR PIECE OF GROUND, situate in the 7th Ward, City of Pittsburgh, County of Allegheny, Pennsylvania, being more particularly described as follows:

BEGINNING on a point on the center line of Bayard Road, said point being located from the intersections of the northeast corner of the Winchester-Thurston School and the northeast side of Morewood Avenue the following courses; thence along the northeast side of Morewood Avenue N 14° 56' 35" W, a distance of 18.24' to a point; thence N 21° 13' 05" W, a distance of 60.22' to the center line of Bayard Road; thence the center line of Bayard Road N 68° 35' 55" E, a distance of 147.00' to the point of beginning; thence continuing along the said center line of Bayard Road N 68° 35' 55" E, a distance of 85.00' to a point of curvature; thence along an arc curving to the left having a radius of 60.00' and length of ~~103.91'~~ **112.05'** to a point; thence N 67° 24' 05" E, a distance of 121.00' to a point; thence N 76° 35' 55" E, a distance of 37.00' to a point of curvature; thence by an arc curving to the left, having a radius of 80.00' and an arc length of 72.61' to a point; thence S 58° 21' 37" E, a distance of 54.82' to a point on the westerly line of Amberson Avenue; thence along the westerly line of Amberson Avenue, S 16° 03' 10" W, a distance of 144.00' to a point; thence continuing along the westerly line of Amberson Avenue S 24° 06' 20" E, a distance of 83.919' to a point of curvature; thence continuing along the westerly line of Amberson Avenue by an arc curving to the right, having a radius of 20.00' and an arc length of 34.63' to a point on the northerly line of Bayard Street; thence along the northerly line of Bayard Street, S 75° 05' 55" W, a distance of 211.50' to a point; thence N 14° 56' 35" W, a distance of 135.00' to a point; thence S 75° 05' 55" W, a distance of 99.77' to a point; thence N 14° 56' 35" W, a distance of 70.63' to a point; thence N 19° 00' 39" W, a distance of ~~87.30'~~ **87.40'** to the center of Bayard Road at the point of beginning.

CONTAINING an area of 79,310.67 Sq./Ft. = 1.8207 acres.

RESERVING unto the declarant for the remaining portion of its property a right of ingress, egress and regress over the private drive from Morewood Avenue as used by the Amberson Towers and Amberson Gardens Apartments and granting to the part herein submitted, a right of egress, ingress and regress over said private right of way as now exists.

The obligation of repairs and maintenance of the said right of way shall be shared equally by all users.

ALSO, reserving for remaining portion of ground of declarant a right for all conduits, telephone, electric, gas and any other utility line, sewer, etc. and granting same over remaining portion of land to property herein submitted.

HAVING erected thereon an apartment building known as 5 Bayard Road.

BEING part of the same property which John A. Cervieri, Jr., et al., by their deed dated December 18, 1976, and recorded in the Recorder's Office of Allegheny County, Pennsylvania, in Deed Book Volume 5721, page 456, granted to First Service Corporation, a Pennsylvania corporation and Amberson Associates, a Pennsylvania Limited Partnership. See Deed Book Volume 5720, page 295, and Deed Book Volume 5723, page 71.

## ARTICLE V {-}

### Section 2 {-}

The proportionate undivided interest in the common elements of each unit is as follows:

| APT. UNIT NO.  | PERCENTAGE | 
| :------------- | :--------- |
| 100            | 0.6747%    |
| 101            | 0.3767%    |
| 102            | 0.3767%    |
| 103            | 0.6747%    |
| 104            | 0.4675%    |
| 105            | 0.4548%    |
| 106            | 0.4633%    |
| 107            | 0.4548%    |
| 109            | 0.6376%    |
| 111            | 0.4570%    |
| 112            | 0.6551%    |
| 113            | 0.6147%    |
| 114            | 0.4675%    |
| 115            | 0.3597%    |
| 116            | 0.4675%    |
| 117            | 0.3442%    |
| 118+120        | 1.0956%    |
| 119            | 0.6747%    |
| 121            | 0.3767%    |
| 200            | 0.6747%    |
| 201            | 0.3767%    |
| 202            | 0.3767%    |
| 203            | 0.6747%    |
| 204            | 0.4675%    |
| 205            | 0.4548%    |
| 206            | 0.4633%    |
| 207            | 0.4548%    |
| 209            | 0.6376%    |
| 210            | 0.4463%    |
| 211            | 0.4570%    |
| 212            | 0.8028%    |
| 213            | 0.6147%    |
| 214            | 0.4675%    |
| 215            | 0.4505%    |
| 216            | 0.4675%    |
| 217            | 0.4548%    |
| 218            | 0.6747%    |
| 219            | 0.6747%    |
| 220            | 0.3767%    |
| 221            | 0.3767%    |
| 300            | 0.6747%    |
| 301            | 0.3767%    |
| 302            | 0.3767%    |
| 303            | 0.6747%    |
| 304            | 0.4675%    |
| 305            | 0.4548%    |
| 306            | 0.4633%    |
| 307            | 0.4548%    |
| 308            | 0.4740%    |
| 309            | 0.6376%    |
| 310            | 0.4463%    |
| 311            | 0.4570%    |
| 312            | 0.8028%    |
| 313            | 0.6147%    |
| 314            | 0.4675%    |
| 315            | 0.4505%    |
| 316            | 0.4675%    |
| 317            | 0.4548%    |
| 318            | 0.6747%    |
| 319            | 0.6747%    |
| 320            | 0.3767%    |
| 321            | 0.3767%    |
| 400            | 0.6747%    |
| 401            | 0.3767%    |
| 402            | 0.3767%    |
| 403            | 0.6747%    |
| 404            | 0.4675%    |
| 405            | 0.4548%    |
| 406            | 0.4633%    |
| 407            | 0.4548%    |
| 408            | 0.4740%    |
| 409            | 0.6376%    |
| 410            | 0.4463%    |
| 411            | 0.4570%    |
| 412            | 0.8028%    |
| 413            | 0.6147%    |
| 414            | 0.4675%    |
| 415            | 0.4505%    |
| 416            | 0.4675%    |
| 417            | 0.4548%    |
| 418            | 0.6747%    |
| 419            | 0.6747%    |
| 420            | 0.3767%    |
| 421            | 0.3767%    |
| 500            | 0.6747%    |
| 501            | 0.3767%    |
| 502            | 0.3767%    |
| 503            | 0.6747%    |
| 504            | 0.4675%    |
| 505            | 0.4548%    |
| 506            | 0.4633%    |
| 507            | 0.4548%    |
| 508            | 0.4740%    |
| 509            | 0.6376%    |
| 510            | 0.4463%    |
| 511            | 0.4570%    |
| 512            | 0.8028%    |
| 513            | 0.6147%    |
| 514            | 0.4675%    |
| 515            | 0.4505%    |
| 516            | 0.4675%    |
| 517            | 0.4548%    |
| 518            | 0.6747%    |
| 519            | 0.6747%    |
| 520            | 0.3767%    |
| 521            | 0.3767%    |
| 600            | 0.6747%    |
| 601            | 0.3767%    |
| 602            | 0.3767%    |
| 603            | 0.6747%    |
| 604            | 0.4675%    |
| 605            | 0.4548%    |
| 606            | 0.4633%    |
| 607            | 0.4548%    |
| 608            | 0.4740%    |
| 609            | 0.6376%    |
| 610            | 0.4463%    |
| 611            | 0.4570%    |
| 612            | 0.8028%    |
| 613            | 0.6147%    |
| 614            | 0.4675%    |
| 615            | 0.4505%    |
| 616            | 0.4675%    |
| 617            | 0.4548%    |
| 618            | 0.6747%    |
| 619            | 0.6747%    |
| 620            | 0.3767%    |
| 621            | 0.3767%    |
| 700            | 0.6747%    |
| 701            | 0.3767%    |
| 702            | 0.3767%    |
| 703            | 0.6747%    |
| 704            | 0.4675%    |
| 705            | 0.4547%    |
| 706            | 0.4633%    |
| 707            | 0.4548%    |
| 708            | 0.4740%    |
| 709            | 0.6376%    |
| 710            | 0.4463%    |
| 711            | 0.4570%    |
| 712            | 0.8028%    |
| 713            | 0.6147%    |
| 714            | 0.4675%    |
| 715            | 0.4505%    |
| 716            | 0.4675%    |
| 717            | 0.4548%    |
| 718            | 0.6747%    |
| 719            | 0.6747%    |
| 720            | 0.3767%    |
| 721            | 0.3767%    |
| 800            | 0.6747%    |
| 801            | 0.3767%    |
| 802            | 0.3767%    |
| 803            | 0.6747%    |
| 804            | 0.4675%    |
| 805            | 0.4548%    |
| 806            | 0.4633%    |
| 807            | 0.4547%    |
| 808            | 0.4740%    |
| 809            | 0.6376%    |
| 810            | 0.4463%    |
| 811            | 0.4570%    |
| 812            | 0.8028%    |
| 813            | 0.6147%    |
| 814            | 0.4675%    |
| 815            | 0.4505%    |
| 816            | 0.4675%    |
| 817            | 0.4548%    |
| 818            | 0.6747%    |
| 819            | 0.6747%    |
| 820            | 0.3767%    |
| 821            | 0.3767%    |
| 900            | 0.6747%    |
| 901            | 0.3767%    |
| 902+903        | 1.0918%    |
| 904            | 0.4675%    |
| 905            | 0.4548%    |
| 906            | 0.4633%    |
| 907            | 0.4548%    |
| 908            | 0.4740%    |
| 909            | 0.6376%    |
| 910            | 0.4463%    |
| 911            | 0.4570%    |
| 912            | 0.8028%    |
| 913            | 0.6147%    |
| 914            | 0.4675%    |
| 915            | 0.4505%    |
| 916            | 0.4675%    |
| 917            | 0.4548%    |
| 918            | 0.6747%    |
| 919+921        | 0.8410%    |
| 920            | 0.3767%    |

## ARTICLE VI {-}

The proportionate undivided interests in the common elements may be altered by the recording of an amendment duly executed by all unit owners affected thereby.

## ARTICLE VII {-}

(A) There shall be no obstruction of any part of the common area. Nothing shall be stored, kept, or parked in the common area without the prior consent of the Council. The use occupancy and operation of the parking garage is under the exclusive control and rules of the Council.

(B) Nothing will be done or kept in any unit or in the common area which will increase the rate of insurance on the building without the prior written consent of Council. No owner shall permit anything to be done or kept in his unit or in the common area which will result in the cancellation of insurance on the building, or which would be in violation of any government statutes, ordinances, rules or regulations. No waste shall be permitted in the common area. Council shall assign suitable locker space to each unit owner, if the same is available.

(C) No unit owner may permit or suffer anything to be done or kept upon the premises which will obstruct or interfere with the rights of other unit owners or annoy other unit owners by unreasonable noise or otherwise, nor which will be noxious or offensive to the other unit owners. Each unit owner shall comply with all of the requirements for all governmental agencies, federal, state, local and all laws, ordinances, rules and regulations applicable to the unit.

## ARTICLE VIII {-}

The names of the first members of the Council are:

1. George Dimeling 
2. Paul Kossis
3. Harry S. Kalson 
4. Albert Raizman 
5. James Kossis

IN WITNESS WHEREOF, the undersigned have hereunto set their hands and seals this 14th day of February, 1978.

FIRST SERVICES CORPORATION  
BY: George Dimeling

AMBERSON ASSOCIATES  
BY: Paul Kosis,  
General Partner

BY: Harry S. Kalson,  
General Partner