### 2022-02-10 tuftetest.tex

\documentclass[]{tufte-handout}

% ams
\usepackage{amssymb,amsmath}

\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \makeatletter
  \@ifpackageloaded{fontspec}{}{\usepackage{fontspec}}
  \makeatother
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
  \makeatletter
  \@ifpackageloaded{soul}{
     \renewcommand\allcapsspacing[1]{{\addfontfeature{LetterSpace=15}#1}}
     \renewcommand\smallcapsspacing[1]{{\addfontfeature{LetterSpace=10}#1}}
   }{}
  \makeatother

\fi

% graphix
\usepackage{graphicx}
\setkeys{Gin}{width=\linewidth,totalheight=\textheight,keepaspectratio}

% booktabs
\usepackage{booktabs}

% url
\usepackage{url}

% hyperref
\usepackage{hyperref}

% units.
\usepackage{units}

\setcounter{secnumdepth}{-1}

% citations
\usepackage{natbib}
\bibliographystyle{plainnat}

% pandoc syntax highlighting

% table with pandoc
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}

% multiplecol
\usepackage{multicol}

% strikeout
\usepackage[normalem]{ulem}

% morefloats
\usepackage{morefloats}

% tightlist macro required by pandoc >= 1.14
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% title / author / date
\title[Tufte Handout with R Markdown]{A Tufte Handout Example}
\author{JJ Allaire and Yihui Xie}
\date{2022-02-10}

\begin{document}

\maketitle

\hypertarget{article-i}{%
\subsection*{ARTICLE I}\label{article-i}}
\addcontentsline{toc}{subsection}{ARTICLE I}

This Declaration is prepared in accordance with the provisions of the
Unit Property Act of the Commonwealth of Pennsylvania (Act of July 3,
1963, P.L. No.~196) for the purpose of submitting to the provisions of
said Act the property described in Article II below.

\hypertarget{article-ii}{%
\subsection*{ARTICLE II}\label{article-ii}}
\addcontentsline{toc}{subsection}{ARTICLE II}

ALL THAT CERTAIN LOT OR PIECE OF GROUND, situate in the 7th Ward, City
of Pittsburgh, County of Allegheny, Pennsylvania, being more
particularly described as follows:

BEGINNING on a point on the center line of Bayard Road, said point being
located from the intersections of the northeast corner of the
Winchester-Thurston School and the northeast side of Morewood Avenue the
following courses; thence along the northeast side of Morewood Avenue N
14° 56' 35'' W, a distance of 18.24' to a point; thence N 21° 13' 05''
W, a distance of 60.22' to the center line of Bayard Road; thence the
center line of Bayard Road N 68° 35' 55'' E, a distance of 147.00' to
the point of beginning; thence continuing along the said center line of
Bayard Road N 68° 35' 55'' E, a distance of 85.00' to a point of
curvature; thence along an arc curving to the left having a radius of
60.00' and length of \sout{103.91'} \textbf{112.05'} to a point; thence
N 67° 24' 05'' E, a distance of 121.00' to a point; thence N 76° 35'
55'' E, a distance of 37.00' to a point of curvature; thence by an arc
curving to the left, having a radius of 80.00' and an arc length of
72.61' to a point; thence S 58° 21' 37'' E, a distance of 54.82' to a
point on the westerly line of Amberson Avenue; thence along the westerly
line of Amberson Avenue, S 16° 03' 10'' W, a distance of 144.00' to a
point; thence continuing along the westerly line of Amberson Avenue S
24° 06' 20'' E, a distance of 83.919' to a point of curvature; thence
continuing along the westerly line of Amberson Avenue by an arc curving
to the right, having a radius of 20.00' and an arc length of 34.63' to a
point on the northerly line of Bayard Street; thence along the northerly
line of Bayard Street, S 75° 05' 55'' W, a distance of 211.50' to a
point; thence N 14° 56' 35'' W, a distance of 135.00' to a point; thence
S 75° 05' 55'' W, a distance of 99.77' to a point; thence N 14° 56' 35''
W, a distance of 70.63' to a point; thence N 19° 00' 39'' W, a distance
of \sout{87.30'} \textbf{87.40'} to the center of Bayard Road at the
point of beginning.

CONTAINING an area of 79,310.67 Sq./Ft. = 1.8207 acres.

RESERVING unto the declarant for the remaining portion of its property a
right of ingress, egress and regress over the private drive from
Morewood Avenue as used by the Amberson Towers and Amberson Gardens
Apartments and granting to the part herein submitted, a right of egress,
ingress and regress over said private right of way as now exists.

The obligation of repairs and maintenance of the said right of way shall
be shared equally by all users.

ALSO, reserving for remaining portion of ground of declarant a right for
all conduits, telephone, electric, gas and any other utility line,
sewer, etc. and granting same over remaining portion of land to property
herein submitted.

HAVING erected thereon an apartment building known as 5 Bayard Road.

BEING part of the same property which John A. Cervieri, Jr., et al., by
their deed dated December 18, 1976, and recorded in the Recorder's
Office of Allegheny County, Pennsylvania, in Deed Book Volume 5721, page
456, granted to First Service Corporation, a Pennsylvania corporation
and Amberson Associates, a Pennsylvania Limited Partnership. See Deed
Book Volume 5720, page 295, and Deed Book Volume 5723, page 71.

\hypertarget{article-v}{%
\subsection*{ARTICLE V}\label{article-v}}
\addcontentsline{toc}{subsection}{ARTICLE V}

\hypertarget{section-2}{%
\subsubsection*{Section 2}\label{section-2}}
\addcontentsline{toc}{subsubsection}{Section 2}

The proportionate undivided interest in the common elements of each unit
is as follows:

\begin{longtable}[]{@{}ll@{}}
\toprule
APT. UNIT NO. & PERCENTAGE \\
\midrule
\endhead
100 & 0.6747\% \\
101 & 0.3767\% \\
102 & 0.3767\% \\
103 & 0.6747\% \\
104 & 0.4675\% \\
105 & 0.4548\% \\
106 & 0.4633\% \\
107 & 0.4548\% \\
109 & 0.6376\% \\
111 & 0.4570\% \\
112 & 0.6551\% \\
113 & 0.6147\% \\
114 & 0.4675\% \\
115 & 0.3597\% \\
116 & 0.4675\% \\
117 & 0.3442\% \\
118+120 & 1.0956\% \\
119 & 0.6747\% \\
121 & 0.3767\% \\
200 & 0.6747\% \\
201 & 0.3767\% \\
202 & 0.3767\% \\
203 & 0.6747\% \\
204 & 0.4675\% \\
205 & 0.4548\% \\
206 & 0.4633\% \\
207 & 0.4548\% \\
209 & 0.6376\% \\
210 & 0.4463\% \\
211 & 0.4570\% \\
212 & 0.8028\% \\
213 & 0.6147\% \\
214 & 0.4675\% \\
215 & 0.4505\% \\
216 & 0.4675\% \\
217 & 0.4548\% \\
218 & 0.6747\% \\
219 & 0.6747\% \\
220 & 0.3767\% \\
221 & 0.3767\% \\
300 & 0.6747\% \\
301 & 0.3767\% \\
302 & 0.3767\% \\
303 & 0.6747\% \\
304 & 0.4675\% \\
305 & 0.4548\% \\
306 & 0.4633\% \\
307 & 0.4548\% \\
308 & 0.4740\% \\
309 & 0.6376\% \\
310 & 0.4463\% \\
311 & 0.4570\% \\
312 & 0.8028\% \\
313 & 0.6147\% \\
314 & 0.4675\% \\
315 & 0.4505\% \\
316 & 0.4675\% \\
317 & 0.4548\% \\
318 & 0.6747\% \\
319 & 0.6747\% \\
320 & 0.3767\% \\
321 & 0.3767\% \\
400 & 0.6747\% \\
401 & 0.3767\% \\
402 & 0.3767\% \\
403 & 0.6747\% \\
404 & 0.4675\% \\
405 & 0.4548\% \\
406 & 0.4633\% \\
407 & 0.4548\% \\
408 & 0.4740\% \\
409 & 0.6376\% \\
410 & 0.4463\% \\
411 & 0.4570\% \\
412 & 0.8028\% \\
413 & 0.6147\% \\
414 & 0.4675\% \\
415 & 0.4505\% \\
416 & 0.4675\% \\
417 & 0.4548\% \\
418 & 0.6747\% \\
419 & 0.6747\% \\
420 & 0.3767\% \\
421 & 0.3767\% \\
500 & 0.6747\% \\
501 & 0.3767\% \\
502 & 0.3767\% \\
503 & 0.6747\% \\
504 & 0.4675\% \\
505 & 0.4548\% \\
506 & 0.4633\% \\
507 & 0.4548\% \\
508 & 0.4740\% \\
509 & 0.6376\% \\
510 & 0.4463\% \\
511 & 0.4570\% \\
512 & 0.8028\% \\
513 & 0.6147\% \\
514 & 0.4675\% \\
515 & 0.4505\% \\
516 & 0.4675\% \\
517 & 0.4548\% \\
518 & 0.6747\% \\
519 & 0.6747\% \\
520 & 0.3767\% \\
521 & 0.3767\% \\
600 & 0.6747\% \\
601 & 0.3767\% \\
602 & 0.3767\% \\
603 & 0.6747\% \\
604 & 0.4675\% \\
605 & 0.4548\% \\
606 & 0.4633\% \\
607 & 0.4548\% \\
608 & 0.4740\% \\
609 & 0.6376\% \\
610 & 0.4463\% \\
611 & 0.4570\% \\
612 & 0.8028\% \\
613 & 0.6147\% \\
614 & 0.4675\% \\
615 & 0.4505\% \\
616 & 0.4675\% \\
617 & 0.4548\% \\
618 & 0.6747\% \\
619 & 0.6747\% \\
620 & 0.3767\% \\
621 & 0.3767\% \\
700 & 0.6747\% \\
701 & 0.3767\% \\
702 & 0.3767\% \\
703 & 0.6747\% \\
704 & 0.4675\% \\
705 & 0.4547\% \\
706 & 0.4633\% \\
707 & 0.4548\% \\
708 & 0.4740\% \\
709 & 0.6376\% \\
710 & 0.4463\% \\
711 & 0.4570\% \\
712 & 0.8028\% \\
713 & 0.6147\% \\
714 & 0.4675\% \\
715 & 0.4505\% \\
716 & 0.4675\% \\
717 & 0.4548\% \\
718 & 0.6747\% \\
719 & 0.6747\% \\
720 & 0.3767\% \\
721 & 0.3767\% \\
800 & 0.6747\% \\
801 & 0.3767\% \\
802 & 0.3767\% \\
803 & 0.6747\% \\
804 & 0.4675\% \\
805 & 0.4548\% \\
806 & 0.4633\% \\
807 & 0.4547\% \\
808 & 0.4740\% \\
809 & 0.6376\% \\
810 & 0.4463\% \\
811 & 0.4570\% \\
812 & 0.8028\% \\
813 & 0.6147\% \\
814 & 0.4675\% \\
815 & 0.4505\% \\
816 & 0.4675\% \\
817 & 0.4548\% \\
818 & 0.6747\% \\
819 & 0.6747\% \\
820 & 0.3767\% \\
821 & 0.3767\% \\
900 & 0.6747\% \\
901 & 0.3767\% \\
902+903 & 1.0918\% \\
904 & 0.4675\% \\
905 & 0.4548\% \\
906 & 0.4633\% \\
907 & 0.4548\% \\
908 & 0.4740\% \\
909 & 0.6376\% \\
910 & 0.4463\% \\
911 & 0.4570\% \\
912 & 0.8028\% \\
913 & 0.6147\% \\
914 & 0.4675\% \\
915 & 0.4505\% \\
916 & 0.4675\% \\
917 & 0.4548\% \\
918 & 0.6747\% \\
919+921 & 0.8410\% \\
920 & 0.3767\% \\
\bottomrule
\end{longtable}

\hypertarget{article-vi}{%
\subsection*{ARTICLE VI}\label{article-vi}}
\addcontentsline{toc}{subsection}{ARTICLE VI}

The proportionate undivided interests in the common elements may be
altered by the recording of an amendment duly executed by all unit
owners affected thereby.

\hypertarget{article-vii}{%
\subsection*{ARTICLE VII}\label{article-vii}}
\addcontentsline{toc}{subsection}{ARTICLE VII}

\begin{enumerate}
\def\labelenumi{(\Alph{enumi})}
\item
  There shall be no obstruction of any part of the common area. Nothing
  shall be stored, kept, or parked in the common area without the prior
  consent of the Council. The use occupancy and operation of the parking
  garage is under the exclusive control and rules of the Council.
\item
  Nothing will be done or kept in any unit or in the common area which
  will increase the rate of insurance on the building without the prior
  written consent of Council. No owner shall permit anything to be done
  or kept in his unit or in the common area which will result in the
  cancellation of insurance on the building, or which would be in
  violation of any government statutes, ordinances, rules or
  regulations. No waste shall be permitted in the common area. Council
  shall assign suitable locker space to each unit owner, if the same is
  available.
\item
  No unit owner may permit or suffer anything to be done or kept upon
  the premises which will obstruct or interfere with the rights of other
  unit owners or annoy other unit owners by unreasonable noise or
  otherwise, nor which will be noxious or offensive to the other unit
  owners. Each unit owner shall comply with all of the requirements for
  all governmental agencies, federal, state, local and all laws,
  ordinances, rules and regulations applicable to the unit.
\end{enumerate}

\hypertarget{article-viii}{%
\subsection*{ARTICLE VIII}\label{article-viii}}
\addcontentsline{toc}{subsection}{ARTICLE VIII}

The names of the first members of the Council are:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  George Dimeling
\item
  Paul Kossis
\item
  Harry S. Kalson
\item
  Albert Raizman
\item
  James Kossis
\end{enumerate}

IN WITNESS WHEREOF, the undersigned have hereunto set their hands and
seals this 14th day of February, 1978.

FIRST SERVICES CORPORATION\\
BY: George Dimeling

AMBERSON ASSOCIATES\\
BY: Paul Kosis,\\
General Partner

BY: Harry S. Kalson,\\
General Partner

\bibliography{skeleton.bib}

\end{document}